Estimating autocorrelation from multiple Markov chains

### 2020-06-11 Paper/main.tex
Proofs clean-up

\documentclass[12pt]{article}
% We can write notes using the percent symbol!
% The first line above is to announce we are beginning a document, an article in this case, and we want the default font size to be 12pt
% \usepackage[utf8]{inputenc}
\usepackage[left=1.2in,top=.8in,right=1.2in,nohead,bottom=.5in]{geometry}
% This is a package to accept utf8 input.  I normally do not use it in my documents, but it was here by default in Overleaf.

\usepackage{amssymb,amsthm,amsmath}
\usepackage{commath}
\usepackage{enumerate,xcolor}
\usepackage{natbib}
\usepackage{graphicx}
\usepackage{caption,setspace}
\usepackage{subcaption}
\DeclareMathOperator{\Tr}{Tr}
\newcommand{\dd}[1]{\mathrm{d}#1}
\newcommand{\E}{\mathbb{E}}
\newcommand{\Var}{\text{Var}}
\newcommand{\X}{\mathcal{X}}
\newcommand{\B}{\mathcal{B}}
% These three packages are from the American Mathematical Society and includes all of the important symbols and operations 
\usepackage{fullpage}
\usepackage{comment}
% By default, an article has some vary large margins to fit the smaller page format.  This allows us to use more standard margins.
\allowdisplaybreaks
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\newtheorem{corollary}{Corollary}

\theoremstyle{remark}
\newtheorem{example}{Example}
\newtheorem{ass}{Assumption}
\newtheorem{remark}{Remark}
\setlength{\parskip}{0em}
% This gives us a full line break when we write a new paragraph
% \usepackage[mathscr]{euscript}
% \usepackage{tikz}
% \newcommand*\circled[1]{\tikz[baseline=(char.base)]{
%             \node[shape=circle,draw,inner sep=2pt] (char) {#1};}}
% \newcommand{\indep}{\perp \!\!\! \perp}

@@

 \begin{equation} \label{eq:priestly}
     \mathbb{E}[\hat{\Gamma}(k) = \left(1- \dfrac{\abs{k}}{n}\right)\left(\Gamma(k) - \Var{(\bar{X})}
 \right)
 \end{equation}

@@

@@

\subsection{Bias Calculations} \label{appendix:bias}

\begin{lemma} \label{lemma:bias1}
$\mathbb{E}\left[ \left(X_{11} - \bar{X}_1 \right) \left(\bar{X}_1 - \bar{\bar{X}} \right)^T \right] = \dfrac{m-1}{mn}\left(\sum\limits_{k=0}^{n-1}\Gamma(k) - \Sigma\right)$
\end{lemma}
\begin{proof}
\begin{align*}
    &\mathbb{E} \left[ \left(X_{11} - \bar{X}_1 \right) \left(\bar{X}_1 - \bar{\bar{X}} \right)^T \right]\\
     &= \mathbb{E} \left[X_{11}\bar{X}_1^T \right] - \mathbb{E}\left[X_{11}\bar{\bar{X}}^T \right] + \mathbb{E} \left[\bar{X}_1\bar{\bar{X}}^T \right] - \mathbb{E} \left[\bar{X}_1\bar{X}_1^T \right]\\
    &= \mathbb{E} \left[X_{11}\bar{X}_1^T \right] - \dfrac{1}{m}\mathbb{E} \left[X_{11}\bar{X}_1^T \right] - \dfrac{m-1}{m}\mathbb{E} \left[X_{11}\bar{X}_2^T \right] + \dfrac{1}{m}\mathbb{E} \left[\bar{X}_1\bar{X}_1^T \right] + \dfrac{m-1}{m}\mathbb{E} \left[\bar{X}_1\bar{X}_2^T \right] - \mathbb{E} \left[\bar{X}_1\bar{X}_1^T \right]\\
    &= \dfrac{m-1}{m}\left(\mathbb{E} \left[X_{11}\bar{X}_1^T \right] - \mathbb{E} \left[X_{11}\bar{X}_2^T \right] + \mathbb{E} \left[\bar{X}_1\bar{X}_2^T \right] - \mathbb{E} \left[\bar{X}_1\bar{X}_1^T \right]\right)\\
    &= \dfrac{m-1}{m}\left(\dfrac{1}{n}\sum_{t=1}^{n}\mathbb{E} \left[X_{11}X_{1t}^T \right] - \mathbb{E}\left[X_{11}\right] \mathbb{E}\left[\bar{X}_2^T \right] + \mathbb{E} \left[\bar{X}_1\right]\mathbb{E}\left[\bar{X}_2^T\right] - \text{Var}\left[\bar{X}_1 \right] - \mathbb{E}\left[\bar{X}_1 \right]\mathbb{E}\left[\bar{X}_1^T \right]\right)\\
    &= \dfrac{m-1}{m}\left(\dfrac{1}{n}\sum_{k=0}^{n-1}\Gamma(k) + \mu \mu^T - \mu \mu^T + \mu \mu^T - \dfrac{\Sigma}{n} - \mu \mu^T\right)\\
    &= \dfrac{m-1}{mn}\left(\sum_{k=0}^{n-1}\Gamma(k) - \Sigma\right)\,.
\end{align*}
%
Similarly, we can show that $\mathbb{E}[A_{ij}^T] = \mathbb{E}[A_{ij}]$ using the symmetry of $\Gamma(k) \textrm{ and } \Sigma$ matrices.
%    
\begin{align*}
    \mathbb{E} \left[ \left(\bar{X}_1 - \bar{\bar{X}} \right)  \left(X_{11} - \bar{X}_1 \right)^T \right] &= \mathbb{E}\left[ \left(X_{11} - \bar{X}_1 \right) \left(\bar{X}_1 - \bar{\bar{X}} \right)^T \right]^T = \dfrac{m-1}{mn}\left(\sum_{k=0}^{n-1}\Gamma(k) - \Sigma \right)^T\,.
    % &= \dfrac{m-1}{mn}\left(\sum_{k=0}^{n-1}\Gamma(k) - \Sigma \right)\\
\end{align*}
\end{proof}

\begin{lemma} \label{lemma:bias2}
\[\mathbb{E}\left[\sum\limits_{s=1}^{m} \left(\bar{X}_{s} - \bar{\bar{X}} \right) \left(\bar{X}_{s} - \bar{\bar{X}} \right)^{T}\right] =  (m-1)\text{Var}(\bar{X}_1)\,.
\]
\end{lemma}

% $\bar{X}_i$ are independently and identically distributed for all $i$; this implies $\Var(\bar{\bar{X}}) = \dfrac{1}{m} \Var(\bar{X}_1)$.

\begin{proof}
\begin{align*}
\mathbb{E}\left[\sum_{s=1}^{m} \left(\bar{X}_{s} - \bar{\bar{X}} \right)  \left(\bar{X}_{s} - \bar{\bar{X}} \right)^{T}\right] &= \mathbb{E}\left[\sum_{s=1}^{m}\bar{X}_{s}\bar{X}_{s}^{T} -\sum_{s=1}^{m}\bar{X}_{s}\bar{\bar{X}}^{T} - \bar{\bar{X}}^{T}\sum_{s=1}^{m}\bar{X}_{s}^{T} + m\bar{\bar{X}}\bar{\bar{X}}^{T}\right]\\
&= \mathbb{E}\left[\sum_{s=1}^{m}\bar{X}_{s}\bar{X}_{s}^{T} - m\bar{\bar{X}}\bar{\bar{X}}^{T}\right]\\
&= m\left(\mathbb{E}\left[\bar{X}_{1}\bar{X}_{1}^{T}\right] - \mathbb{E}\left[\bar{\bar{X}}\bar{\bar{X}}^{T}\right]\right)\\
&= m\left(\Var(\bar{X}_{1}) + \mu\mu^{T} - \Var(\bar{\bar{X}}) - \mu\mu^{T}\right)\\
&= (m-1)\Var(\overline{X_1})\,.
\end{align*}
\end{proof}

\begin{proof}[Proof of Theorem~\ref{th:RAC_expec}]
\begin{align*}
    & \hat{\Gamma}_{RAC}(k) \\
    &= \dfrac{1}{m}\sum_{s=1}^{m}\hat{\Gamma}_{RAC,s}(k) = \dfrac{1}{mn}\sum_{s=1}^{m}\sum_{t=1}^{n-|k|} \left(X_{st} - \bar{\bar{X}} \right) \left(X_{s(t+k)} - \bar{\bar{X}} \right)^T\\
    &= \left[\dfrac{1}{mn}\sum_{s=1}^{m}\sum_{t=1}^{n-|k|} \left(X_{st} - \bar{X}_s \right) \left(X_{s(t+k)} - \bar{X}_s \right)^T \right] + \left[\dfrac{1}{mn}\sum_{s=1}^{m}\sum_{t=1}^{|k|} \left(\bar{X}_s - \bar{\bar{X}} \right)  \left(\bar{X}_s - X_{st} \right)^T\right]\\ 
    & \quad +  \left[\dfrac{1}{mn}\sum_{s=1}^{m}\sum_{t=n-|k|+1}^{n}  \left( \bar{X}_s - X_{st} \right)  \left(\bar{X}_s - \bar{\bar{X}} \right)^T\right] + \left[\dfrac{n-|k|}{mn}\sum_{s=1}^{m}  \left(\bar{X}_s - \bar{\bar{X}}\right)   \left(\bar{X}_s - \bar{\bar{X}} \right)^T\right]\\
    &= \dfrac{1}{m}\sum_{s=1}^{m}\hat{\Gamma}_s(k) - \dfrac{1}{mn}\sum_{s=1}^{m}\sum_{t=1}^{|k|}A_{st}^T - \dfrac{1}{mn}\sum_{s=1}^{m}\sum_{t=n-|k|+1}^{n}A_{st} + \dfrac{n-|k|}{mn}\sum_{s=1}^{m}  \left(\bar{X}_s - \bar{\bar{X}} \right)  \left(\bar{X}_s - \bar{\bar{X}} \right)^T\,,
\end{align*}
%
where $A_{st} = (X_{st}-\bar{X}_s)(\bar{X}_s - \bar{\bar{X}})^T$. Using \eqref{eq:priestly}, Lemma~\ref{lemma:bias1}, and Lemma~\ref{lemma:bias2}, the expectation of $\hat{\Gamma}_{RAC}(k)$ can be written as 
%
\begin{align*}
\hat{\Gamma}_{RAC}(k) &= \left(1- \dfrac{|k|}{n}\right) \left(\Gamma(k) - \Var(\overline{X_1})\right) - \dfrac{1}{n} \left(\sum\limits_{t=1}^{|k|}\mathbb{E}[A_{1t}^T] + \sum\limits_{t=n-|k|+1}^{n}\mathbb{E}[A_{1t}]\right)\\
    & \quad  + \left(1- \dfrac{|k|}{n}\right)\left(1-\dfrac{1}{m}\right)\Var(\bar{X}_1)\\
    &= \left(1- \dfrac{|k|}{n}\right)\left(\Gamma(k) - \dfrac{1}{m}\Var(\bar{X}_1)\right) - \dfrac{2|k|}{n}\dfrac{m-1}{mn}\left(\sum_{h=0}^{n-1}\Gamma(h) - \Sigma\right)
\end{align*}
\\
Using the results of \cite{song1995optimal} given in \eqref{eq:song} to expand $\Var(\bar{X}_1)$,
\[
\mathbb{E}\left[\hat{\Gamma}_{RAC}(k)\right] = \left(1- \dfrac{|k|}{n}\right)\left(\Gamma(k) - \dfrac{\Sigma}{mn} - \dfrac{\Phi}{mn^2}\right)  + \dfrac{2(m-1)}{m}\dfrac{\abs{k}}{n^2} \left(\Sigma - \sum_{h=0}^{n-1}\Gamma(h)\right) + o(n^{-2})\,.
\]
%
We observe that this expectation result is very similar to equation \ref{eq:priestly} from \cite{priestley1981spectral} for the asymptotically unbiased estimator of autocovariance except for an additional positive bias term. This helps in countering the inherent negative bias in $\hat{\Gamma}(k)$ and give more informative estimates of autocovariance.  Although, bias term for a fixed $k$ is $o(1)$, it increases with $k$. We have explicitly written the end effect term arising from using the estimator $\bar{\bar{X}}$ for $\mu$. These are small order terms that vanish as $n \to \infty$.
\end{proof}

\begin{align*}
    \hat{\Gamma}_{RAC}(k) &= \dfrac{1}{m}\sum_{s=1}^{m}\hat{\Gamma}_{RAC,s}(k) = \dfrac{1}{mn}\sum_{s=1}^{m}\sum_{t=1}^{n}  \left(X_{st} - \bar{\bar{X}}\right)  \left(X_{s(t+k)} - \bar{\bar{X}}\right)^T\\
    &= \left[\dfrac{1}{mn}\sum_{s=1}^{m}\sum_{t=1}^{n} \left(X_{st} - \bar{X}_s \right)  \left(X_{s(t+k)} - \bar{X}_s \right)^T\right] + \left[\dfrac{1}{m}\sum_{s=1}^{m} \left(\bar{X}_s - \bar{\bar{X}} \right) \left(\bar{X}_s - \bar{\bar{X}} \right)^T\right]\\
    &= \dfrac{1}{m}\sum_{s=1}^{m}\hat{\Gamma}_s(k)  + \dfrac{1}{m}\sum_{s=1}^{m} \left(\bar{X}_s - \bar{\bar{X}} \right)  \left(\bar{X}_s - \bar{\bar{X}} \right)^T\,.
\end{align*}
Using \eqref{eq:priestly}, Lemma~\ref{lemma:bias2}, and \eqref{eq:song} the expectation of $\hat{\Gamma}_{RAC}(0)$ can be written as 
%
\begin{align*}
    \mathbb{E} \left[\hat{\Gamma}_{RAC}(0) \right] &= \left(\Gamma(k) - \Var(\bar{X}_1)\right) + \left(1-\dfrac{1}{m}\right)\Var(\bar{X}_1) = \Gamma(k) - \dfrac{1}{m}\Var(\bar{X}_1)\\
    &= \Gamma(0) - \dfrac{\Sigma}{mn} - \dfrac{\Phi}{mn^2} + o(n^{-2})
\end{align*}

>>>>

\documentclass[11pt]{article}
% We can write notes using the percent symbol!
% The first line above is to announce we are beginning a document, an article in this case, and we want the default font size to be 12pt
% \usepackage[utf8]{inputenc}
\usepackage[left=1.2in,top=.8in,right=1.2in,nohead,bottom=.5in]{geometry}
% This is a package to accept utf8 input.  I normally do not use it in my documents, but it was here by default in Overleaf.

\usepackage{amssymb,amsthm,amsmath}
\usepackage{commath}
\usepackage{enumerate,xcolor}
\usepackage{natbib}
\usepackage{graphicx}
\usepackage{caption,setspace}
\usepackage{subcaption}
\DeclareMathOperator{\Tr}{Tr}
\newcommand{\dd}[1]{\mathrm{d}#1}
\newcommand{\E}{\mathbb{E}}
\newcommand{\Var}{\text{Var}}
\newcommand{\X}{\mathcal{X}}
\newcommand{\B}{\mathcal{B}}
\newcommand\numberthis{\addtocounter{equation}{1}\tag{\theequation}}
% These three packages are from the American Mathematical Society and includes all of the important symbols and operations 
\usepackage{fullpage}
\usepackage{comment}
% By default, an article has some vary large margins to fit the smaller page format.  This allows us to use more standard margins.
\allowdisplaybreaks
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\newtheorem{corollary}{Corollary}

\theoremstyle{remark}
\newtheorem{example}{Example}
\newtheorem{ass}{Assumption}
\newtheorem{remark}{Remark}
\setlength{\parskip}{0em}
\setlength{\parindent}{0in}
\renewcommand{\arraystretch}{1.5}

% This gives us a full line break when we write a new paragraph
% \usepackage[mathscr]{euscript}
% \usepackage{tikz}
% \newcommand*\circled[1]{\tikz[baseline=(char.base)]{
%             \node[shape=circle,draw,inner sep=2pt] (char) {#1};}}
% \newcommand{\indep}{\perp \!\!\! \perp}

@@

 \begin{equation} \label{eq:priestly}
     \mathbb{E}[\hat{\Gamma}(k)] = \left(1- \dfrac{\abs{k}}{n}\right)\left(\Gamma(k) - \Var{(\bar{X})}
 \right)
 \end{equation}

@@

\subsection{Bias of autocovariance} \label{appendix:bias}

% \begin{lemma} \label{lemma:bias1}
% $\mathbb{E}\left[ \left(X_{11} - \bar{X}_1 \right) \left(\bar{X}_1 - \bar{\bar{X}} \right)^T \right] = \dfrac{m-1}{mn}\left(\sum\limits_{k=0}^{n-1}\Gamma(k) - \Sigma\right)$
% \end{lemma}
% \begin{proof}
% \begin{align*}
%     &\mathbb{E} \left[ \left(X_{11} - \bar{X}_1 \right) \left(\bar{X}_1 - \bar{\bar{X}} \right)^T \right]\\
%      &= \mathbb{E} \left[X_{11}\bar{X}_1^T \right] - \mathbb{E}\left[X_{11}\bar{\bar{X}}^T \right] + \mathbb{E} \left[\bar{X}_1\bar{\bar{X}}^T \right] - \mathbb{E} \left[\bar{X}_1\bar{X}_1^T \right]\\
%     &= \mathbb{E} \left[X_{11}\bar{X}_1^T \right] - \dfrac{1}{m}\mathbb{E} \left[X_{11}\bar{X}_1^T \right] - \dfrac{m-1}{m}\mathbb{E} \left[X_{11}\bar{X}_2^T \right] + \dfrac{1}{m}\mathbb{E} \left[\bar{X}_1\bar{X}_1^T \right] + \dfrac{m-1}{m}\mathbb{E} \left[\bar{X}_1\bar{X}_2^T \right] - \mathbb{E} \left[\bar{X}_1\bar{X}_1^T \right]\\
%     &= \dfrac{m-1}{m}\left(\mathbb{E} \left[X_{11}\bar{X}_1^T \right] - \mathbb{E} \left[X_{11}\bar{X}_2^T \right] + \mathbb{E} \left[\bar{X}_1\bar{X}_2^T \right] - \mathbb{E} \left[\bar{X}_1\bar{X}_1^T \right]\right)\\
%     &= \dfrac{m-1}{m}\left(\dfrac{1}{n}\sum_{t=1}^{n}\mathbb{E} \left[X_{11}X_{1t}^T \right] - \mathbb{E}\left[X_{11}\right] \mathbb{E}\left[\bar{X}_2^T \right] + \mathbb{E} \left[\bar{X}_1\right]\mathbb{E}\left[\bar{X}_2^T\right] - \text{Var}\left[\bar{X}_1 \right] - \mathbb{E}\left[\bar{X}_1 \right]\mathbb{E}\left[\bar{X}_1^T \right]\right)\\
%     &= \dfrac{m-1}{m}\left(\dfrac{1}{n}\sum_{k=0}^{n-1}\Gamma(k) + \mu \mu^T - \mu \mu^T + \mu \mu^T - \text{Var}\left[\bar{X}_1 \right] - \mu \mu^T\right)\\
%     &= \dfrac{m-1}{mn}\left(\sum_{k=0}^{n-1}\Gamma(k) - n\text{Var}\left[\bar{X}_1 \right]\right)\,.
% \end{align*}
% %
% Similarly,
% %    
% \begin{align*}
%     \mathbb{E} \left[ \left(\bar{X}_1 - \bar{\bar{X}} \right)  \left(X_{11} - \bar{X}_1 \right)^T \right] &= \mathbb{E}\left[ \left(X_{11} - \bar{X}_1 \right) \left(\bar{X}_1 - \bar{\bar{X}} \right)^T \right]^T = \dfrac{m-1}{mn}\left(\sum_{k=0}^{n-1}\Gamma(k) - \Sigma \right)^T\,.
%     % &= \dfrac{m-1}{mn}\left(\sum_{k=0}^{n-1}\Gamma(k) - \Sigma \right)\\
% \end{align*}
% \end{proof}

% \begin{lemma} \label{lemma:bias2}
% \[
% \mathbb{E}\left[\sum\limits_{s=1}^{m} \left(\bar{X}_{s} - \bar{\bar{X}} \right) \left(\bar{X}_{s} - \bar{\bar{X}} \right)^{T}\right] =  (m-1)\text{Var}(\bar{X}_1)\,.
% \]
% \end{lemma}

% % $\bar{X}_i$ are independently and identically distributed for all $i$; this implies $\Var(\bar{\bar{X}}) = \dfrac{1}{m} \Var(\bar{X}_1)$.

% \begin{proof}
% \begin{align*}
% \mathbb{E}\left[\sum_{s=1}^{m} \left(\bar{X}_{s} - \bar{\bar{X}} \right)  \left(\bar{X}_{s} - \bar{\bar{X}} \right)^{T}\right] &= \mathbb{E}\left[\sum_{s=1}^{m}\bar{X}_{s}\bar{X}_{s}^{T} -\sum_{s=1}^{m}\bar{X}_{s}\bar{\bar{X}}^{T} - \bar{\bar{X}}^{T}\sum_{s=1}^{m}\bar{X}_{s}^{T} + m\bar{\bar{X}}\bar{\bar{X}}^{T}\right]\\
% &= \mathbb{E}\left[\sum_{s=1}^{m}\bar{X}_{s}\bar{X}_{s}^{T} - m\bar{\bar{X}}\bar{\bar{X}}^{T}\right]\\
% &= m\left(\mathbb{E}\left[\bar{X}_{1}\bar{X}_{1}^{T}\right] - \mathbb{E}\left[\bar{\bar{X}}\bar{\bar{X}}^{T}\right]\right)\\
% &= m\left(\Var(\bar{X}_{1}) + \mu\mu^{T} - \Var(\bar{\bar{X}}) - \mu\mu^{T}\right)\\
% &= (m-1)\Var(\overline{X_1})\,.
% \end{align*}
% \end{proof}

\begin{proof}[Proof of Theorem~\ref{th:RAC_expec}]
\begin{align*}
    & \hat{\Gamma}_{RAC,s}(k) \\
   & = \dfrac{1}{n}\sum_{t=1}^{n-|k|} \left(X_{st} - \bar{\bar{X}} \right) \left(X_{s(t+k)} - \bar{\bar{X}} \right)^T \\
    &= \left[\dfrac{1}{n} \sum_{t=1}^{n-|k|} \left(X_{st} - \bar{X}_s \right) \left(X_{s(t+k)} - \bar{X}_s \right)^T \right] + \left[\dfrac{1}{n} \sum_{t=1}^{|k|} \left(\bar{X}_s - \bar{\bar{X}} \right)  \left(\bar{X}_s - X_{st} \right)^T\right] \\ 
    & \quad +  \left[\dfrac{1}{n} \sum_{t=n-|k|+1}^{n}  \left( \bar{X}_s - X_{st} \right)  \left(\bar{X}_s - \bar{\bar{X}} \right)^T\right] + \left[\dfrac{n-|k|}{n} \left(\bar{X}_s - \bar{\bar{X}}\right)   \left(\bar{X}_s - \bar{\bar{X}} \right)^T\right]\\
    &= \hat{\Gamma}_s(k) - \dfrac{1}{n} \sum_{t=1}^{|k|}A_{st}^T - \dfrac{1}{n}\sum_{t=n-|k|+1}^{n}A_{st} + \dfrac{n-|k|}{n}  \left(\bar{X}_s - \bar{\bar{X}} \right)  \left(\bar{X}_s - \bar{\bar{X}} \right)^T\,, \numberthis \label{eq:acf_breakdown}
\end{align*}
%
where $A_{st} = (X_{st}-\bar{X}_s)(\bar{X}_s - \bar{\bar{X}})^T$. We will study the expectations of the each of the above terms. Without loss of generality, consider $A_{11}$,
\begin{align*}
    &\E \left[A_{11} \right]\\
     &= \mathbb{E} \left[ \left(X_{11} - \bar{X}_1 \right) \left(\bar{X}_1 - \bar{\bar{X}} \right)^T \right]\\
     &= \mathbb{E} \left[X_{11}\bar{X}_1^T \right] - \mathbb{E}\left[X_{11}\bar{\bar{X}}^T \right] + \mathbb{E} \left[\bar{X}_1\bar{\bar{X}}^T \right] - \mathbb{E} \left[\bar{X}_1\bar{X}_1^T \right]\\
    &= \mathbb{E} \left[X_{11}\bar{X}_1^T \right] - \dfrac{1}{m}\mathbb{E} \left[X_{11}\bar{X}_1^T \right] - \dfrac{m-1}{m}\mathbb{E} \left[X_{11}\bar{X}_2^T \right] + \dfrac{1}{m}\mathbb{E} \left[\bar{X}_1\bar{X}_1^T \right] + \dfrac{m-1}{m}\mathbb{E} \left[\bar{X}_1\bar{X}_2^T \right] - \mathbb{E} \left[\bar{X}_1\bar{X}_1^T \right]\\
    &= \dfrac{m-1}{m}\left(\mathbb{E} \left[X_{11}\bar{X}_1^T \right] - \mathbb{E} \left[X_{11}\bar{X}_2^T \right] + \mathbb{E} \left[\bar{X}_1\bar{X}_2^T \right] - \mathbb{E} \left[\bar{X}_1\bar{X}_1^T \right]\right)\\
    &= \dfrac{m-1}{m}\left(\dfrac{1}{n}\sum_{t=1}^{n}\mathbb{E} \left[X_{11}X_{1t}^T \right] - \mathbb{E}\left[X_{11}\right] \mathbb{E}\left[\bar{X}_2^T \right] + \mathbb{E} \left[\bar{X}_1\right]\mathbb{E}\left[\bar{X}_2^T\right] - \text{Var}\left[\bar{X}_1 \right] - \mathbb{E}\left[\bar{X}_1 \right]\mathbb{E}\left[\bar{X}_1^T \right]\right)\\
    &= \dfrac{m-1}{m}\left(\dfrac{1}{n}\sum_{k=0}^{n-1}\Gamma(k) + \mu \mu^T - \mu \mu^T + \mu \mu^T - \text{Var}\left[\bar{X}_1 \right] - \mu \mu^T\right)\\
    &= \dfrac{m-1}{mn}\left(\sum_{k=0}^{n-1}\Gamma(k) - n\text{Var}\left[\bar{X}_1 \right]\right)\,. \numberthis \label{eq:acf_2}
\end{align*}
%
Similarly,
%    
\begin{align}
\label{eq:acf_3}
    \mathbb{E} \left[ A_{11}^T \right] &= \mathbb{E}\left[ A_{11}\right]^T = \dfrac{m-1}{mn}\left(\sum_{k=0}^{n-1}\Gamma(k) - n\text{Var}\left[\bar{X}_1 \right] \right)^T\,.
    % &= \dfrac{m-1}{mn}\left(\sum_{k=0}^{n-1}\Gamma(k) - \Sigma \right)\\
\end{align}

Further,
\begin{align*}
\mathbb{E}\left[ \left(\bar{X}_{s} - \bar{\bar{X}} \right)  \left(\bar{X}_{s} - \bar{\bar{X}} \right)^{T}\right] &= \mathbb{E}\left[ \bar{X}_{s}\bar{X}_{s}^{T} - \bar{X}_{s}\bar{\bar{X}}^{T} - \bar{\bar{X}} \bar{X}_{s}^{T} + \bar{\bar{X}}\bar{\bar{X}}^{T}\right]\\
&= \mathbb{E}\left[ \bar{X}_{s}\bar{X}_{s}^{T} - \bar{\bar{X}}\bar{\bar{X}}^{T}\right]\\
&= \left(\mathbb{E}\left[\bar{X}_{1}\bar{X}_{1}^{T}\right] - \mathbb{E}\left[\bar{\bar{X}}\bar{\bar{X}}^{T}\right]\right)\\
&= \left(\Var(\bar{X}_{1}) + \mu\mu^{T} - \Var(\bar{\bar{X}}) - \mu\mu^{T}\right)\\
&= \dfrac{m-1}{m}\Var(\bar{X}_1)\,. \numberthis \label{eq:acf_4}
\end{align*}
%
Using \eqref{eq:priestly}, \eqref{eq:acf_2}, \eqref{eq:acf_3}, and \eqref{eq:acf_4} in \eqref{eq:acf_breakdown},
\begin{align*}
& \E \left[\hat{\Gamma}_{RAC,s}(k) \right] \\
&= \mathbb{E}\left[\hat{\Gamma}_{s}(k)\right] - \dfrac{1}{n} \left(\sum\limits_{t=1}^{|k|}\mathbb{E}[A_{1t}^T] + \sum\limits_{t=n-|k|+1}^{n}\mathbb{E}[A_{1t}]\right) + \left(1- \dfrac{|k|}{n}\right)\left(1-\dfrac{1}{m}\right)\Var(\bar{X}_1)\\
    &= \mathbb{E}\left[\hat{\Gamma}_{s}(k)\right] - {\color{red}\dfrac{2|k|}{n}\dfrac{m-1}{mn}\left(\sum_{h=0}^{n-1}\Gamma(h) - n \Var(\bar{X}_1)\right)} + \left(1- \dfrac{|k|}{n}\right)\left(1-\dfrac{1}{m}\right)\Var(\bar{X}_1)\,.
\end{align*}
{\color{red} DV: Instead of the ``2'', should it be something that has $\Gamma(h) + \Gamma(h)^T$?}

Using the results of \cite{song1995optimal} given in \eqref{eq:song} to expand $\Var(\bar{X}_1)$,
\[
\mathbb{E}\left[\hat{\Gamma}_{RAC}(k)\right] = \left(1- \dfrac{|k|}{n}\right)\left(\Gamma(k) - \dfrac{\Sigma}{mn} - \dfrac{\Phi}{mn^2}\right)  + \dfrac{2(m-1)}{m}\dfrac{\abs{k}}{n^2} \left(\Sigma - \sum_{h=0}^{n-1}\Gamma(h)\right) + o(n^{-2})\,.
\]
%
We observe that this expectation result is very similar to equation \ref{eq:priestly} from \cite{priestley1981spectral} for the asymptotically unbiased estimator of autocovariance except for an additional positive bias term. This helps in countering the inherent negative bias in $\hat{\Gamma}(k)$ and give more informative estimates of autocovariance.  Although, bias term for a fixed $k$ is $o(1)$, it increases with $k$. We have explicitly written the end effect term arising from using the estimator $\bar{\bar{X}}$ for $\mu$. These are small order terms that vanish as $n \to \infty$.
\end{proof}

\begin{align*}
    \hat{\Gamma}_{RAC}(0) &= \dfrac{1}{m}\sum_{s=1}^{m}\hat{\Gamma}_{RAC,s}(0) = \dfrac{1}{mn}\sum_{s=1}^{m}\sum_{t=1}^{n}  \left(X_{st} - \bar{\bar{X}}\right)  \left(X_{st} - \bar{\bar{X}}\right)^T\\
    &= \left[\dfrac{1}{mn}\sum_{s=1}^{m}\sum_{t=1}^{n} \left(X_{st} - \bar{X}_s \right)  \left(X_{st} - \bar{X}_s \right)^T\right] + \left[\dfrac{1}{m}\sum_{s=1}^{m} \left(\bar{X}_s - \bar{\bar{X}} \right) \left(\bar{X}_s - \bar{\bar{X}} \right)^T\right]\\
    &= \dfrac{1}{m}\sum_{s=1}^{m}\hat{\Gamma}_s(k)  + \dfrac{1}{m}\sum_{s=1}^{m} \left(\bar{X}_s - \bar{\bar{X}} \right)  \left(\bar{X}_s - \bar{\bar{X}} \right)^T\,.
\end{align*}
Using \eqref{eq:priestly}, Lemma~\ref{lemma:bias2}, and \eqref{eq:song} the expectation of $\hat{\Gamma}_{RAC}(0)$ can be written as 
%
\begin{align*}
    \mathbb{E} \left[\hat{\Gamma}_{RAC}(0) \right] &= \left(\Gamma(k) - \Var(\bar{X}_1)\right) + \left(1-\dfrac{1}{m}\right)\Var(\bar{X}_1) = \Gamma(k) - \dfrac{1}{m}\Var(\bar{X}_1)\\
    &= \Gamma(0) - \dfrac{\Sigma}{mn} - \dfrac{\Phi}{mn^2} + o(n^{-2})
\end{align*}

@@

\subsection{RSV Bias}
